Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-12-02 Appendix.tex
Overview

>>>>

\chapter*{MATLAB Code}
\addcontentsline{toc}{section}{MATLAB Code}
\section*{Example code for Timoshenko beam model}
\lstinputlisting{code/timoshenko_beam/TimoshenkoEig.m}

\section*{Example code for two-dimensional elastic body using bi-cubics}
\lstinputlisting{code/two_dimensional_model/TwoDimensionalCantileverCubic.m}

\section*{Example code for Reissner-Mindlin plate model using bi-cubics}
\lstinputlisting{code/plate_model/PlateCantileverCubic.m}

\section*{Example code for three-dimensional elastic body using tri-cubics}
\lstinputlisting{code/three_dimensional_model/Copy_of_CubePC.m}

### 2023-12-02 files/Overview.tex
Overview

Section 4.5 is a discussion of the article \cite{LVV09}. In this article, the authors investigate the validity of a cantilever Timoshenko beam model, by comparing it to a two-dimensional cantilever Timoshenko beam model. The authors compare the models by comparing the eigenvalues and eigenfunctions. (See Chapter 2 on modal analysis). The two-dimensional model is more complex and there are eigenvalues that are not shared between the models. The authors use the mode shapes to match up the eigenvalues. The eigenvalues matching the eigenvalues of the Timoshenko beam model are referred to by the authors as beam-type eigenvalues. The authors also consider different shapes of beams, from a short thick beam to a long slender beam. The results show that the models compare very well, even for a short and thick beam. \textcolor{red}{The results improve as the beam gets more slender. This shows that the shape of the beam is important. The cases covered are realistic to applications.}\sout{As the beam gets more slender, the comparison improves. In this section some of the results of the article are discussed and replicated.}

The next section is a discussion on the article \cite{SP06}. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors report on an experiment with forced vibration a free-free beam where the natural frequencies are measured. These empirical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. \sout{Finally the chapter is concluded by looking at a model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. A similar model using realistic two- or three-dimensional model will be very challenging.}

@@

\subsubsection{Chapter 6}
\textcolor{red}{This chapter is an extension of the work of Section 4.5. In Section 4.5, the validity of the Timoshenko beam theory is investigated by comparing a Timoshenko beam model to a two-dimensional beam model.}

@@

\sout{In this chapter, the main comparisons of the models are conducted. All of the sections follow the same idea and struture as the article \cite{LVV09}.}

\sout{The first section is a discussion of the article \cite{LVV09}. In this article, the authors investigate the validity of a cantilever Timoshenko beam model, by comparing it to a two-dimensional cantilever Timoshenko beam model. The authors compare the models by comparing the eigenvalues and eigenfunctions (see Chapter 2 on modal analysis). The two-dimensional model is more complex and there are eigenvalues that are not shared between the models. The authors use the mode shapes to match up the eigenvalues. The eigenvalues matching the eigenvalues of the Timoshenko beam model are referred to by the authors as beam-type eigenvalues. The authors also consider different shapes of beams, from a short thick beam to a long slender beam. The results show that the models compare very well, even for a short and thick beam. As the beam gets more slender, the comparison improves. In this section some of the results of the article are discussed and replicated.}

>>>>

Section 4.5 is a discussion of the article \cite{LVV09}. In this article, the authors investigate the validity of a cantilever Timoshenko beam model, by comparing it to a two-dimensional cantilever Timoshenko beam model. The authors compare the models by comparing the eigenvalues and eigenfunctions. (See Chapter 2 on modal analysis). The two-dimensional model is more complex and there are eigenvalues that are not shared between the models. The authors use the mode shapes to match up the eigenvalues. The eigenvalues matching the eigenvalues of the Timoshenko beam model are referred to by the authors as beam-type eigenvalues. The authors also consider different shapes of beams, from a short thick beam to a long slender beam. The results show that the models compare very well, even for a short and thick beam.

The next section is a discussion on the article \cite{SP06}. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors report on an experiment with forced vibration a free-free beam where the natural frequencies are measured. These empirical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory.

@@

\subsubsection{Chapter 6}
This chapter is an extension of the work of Section 4.5. In Section 4.5, the validity of the Timoshenko beam theory is investigated by comparing a Timoshenko beam model to a two-dimensional beam model.

@@

The same method is used in this chapter as in Section 4.5. The mode shapes are sketched and matched. The corresponding eigenvalues can then be matched up and compared. The eigenvalues relating to the Reissner-Mindlin plate model are referred to as plate-type eigenvalues.

\section{Achievements}

In the first chapter, the models of this dissertation is presented and a dimensionless variational form is derived. The model problems used in the dissertation is also presented.

In the second chapter, we look at the existence and uniqueness of solutions for a general vibration problem. All of the models in this dissertation is a special case of this general vibration problem. So a cantilever Timoshenko beam model is used as an example of the application of the theory as well as to explain the theory. A weak variational form of a cantilever Timoshenko beam model is derived. The authors provide four assumptions, and using these assumptions they prove using semi-group theory that the general vibration problem has a unique solution. For the example, we show that the four assumptions can be proven for the cantilever Timoshenko beam model.

Then we look at modal analysis. Again we use the cantilever Timoshenko beam model as an example. A trial solution is chosen and using this trial solution, two problems are obtained. The first is the eigenvalue problem and the second is an ordinary differential equation. These two problems are solved and substitution then shows that the trial solution is valid. This is similar to the separation of variables, but it avoids any division. We then discuss a general solution using the same idea. The idea of modal analysis is crucial to the dissertation. From modal analysis, we can compare the eigenvalues and eigenfunctions of the models, and if they compare well, then the solutions of the models will compare well. And since the formal series solution is a linear combination of Fourier Coefficients, we only need to consider the first few eigenvalues.

In the third chapter, two theoretical results are considered for the Finite Element Method. The first results is on the convergence of the Galerkin Approximation of a general vibration problem. In this section, the results of the article is discussed. The results are also given in Theorems with simple notation. We then look at a result form a textbook on the convergence of the eigenvalues and eigenfunctions when using the Finite Element Method. We discuss the results of the textbook, with some improvements to the notation as well as expanding some results to make it more clear.

In the fourth chapter, we look at our simplest model, the Timoshenko beam model. The first result discusses an article that provides a method to determine the exact eigenvalues and eigenfunctions for the Timoshenko beam model. We then apply this to a cantilever beam model and a free-free beam model to obtain the eigenvalues and plot the mode shapes. We then look at an article that compares a cantilever Timoshenko beam model to a two-dimensional beam model. We discuss the results of the article and replicate the results with more significant digits. We also include some more results not included in the article, that are used later in Chapter 6. We then look at another article that compares the Timoshenko beam theory to results obtained from a measured beam. The authors suspended a beam on both ends an vibrated the beam until then endpoints could be considered free. They then swept through different frequencies and recorded the natural frequencies of the beam. These results are then compared to the theoretical results of a free-free Timoshenko beam. We add to the article by giving the model problems for the beams and discuss the results.

The fifth chapter is a preparation chapter for the comparisons of this dissertation. In chapter 4, we have a method to obtain the exact eigenvalues and eigenfunctions for the Timoshenko beam theory. But a different approach is required for the two- and three-dimensional models. In this chapter, we derive the Galerkin Approximation for each of the models. We then define a discrete grid of elements on the reference configuration on each of the models. Then using piece-wise Hermite cubic basis functions, we derive a matrix representation of the model problems using the standard Finite Element Analysis. In this form it is then easy to obtain a matrix representation of the eigenvalue problem for each of the models.

\section{Future Work}
